---
title: "LinearRegression"
author: "Ethan Michaels"
date: "2024-08-13"
output: html_document
---
```{r}
library(dplyr)
library(tidyverse)
library(randomForest)
library(caret)
library(car)
library(infotheo)
library(pls)
library(ggplot2)
library(olsrr)
library(corrr)
library(glmnet)
library(Metrics)
```

```{r}
MVPMLB <- read.csv("MVPALNLContendersModified.csv", header = TRUE)
Batting2023 <- read.csv("BattingData2.csv", header = TRUE)
```

```{r}
Batting2023$BA <- Batting2023$H/Batting2023$AB
Batting2023AL <- subset(Batting2023,League=="AL")
Batting2023NL <- subset(Batting2023,League=="NL")
```

```{r}
MVPAL <- MVPMLB[MVPMLB$League=="AL",]
MVPNL <- MVPMLB[MVPMLB$League=="NL",]
MVPWinners <- MVPMLB[MVPMLB$Winners==1,]
MVPLosers<- MVPMLB[MVPMLB$Winners==0,]
```

#Descriptive Statistics 
```{r}
teamMVP <- MVPWinners %>% group_by(Tm) %>% count()
teamContenders <- MVPMLB %>% group_by(Tm) %>% count()

```

```{r}
summary(MVPNL)
```

```{r}
summary(MVPAL)

```

```{r}
summary(MVPMLB)

```
```{r}
summary(MVPWinners)
```

```{r}
summary(MVPLosers)
```


#Correlation on Share 
```{r}
cor(MVPMLB[,c('Share','WAR','G','AB','R','H','HR','RBI','SB','BB','BA','OBP','SLG','OPS')])
```

```{r}
cor(MVPAL[,c('Share','WAR','G','AB','R','H','HR','RBI','SB','BB','BA','OBP','SLG','OPS')])
```

```{r}
MVPMLB %>%
  correlate() %>%
  focus(Share)
```

```{r}
#mutinformation(discretize(MVPAL[,8]),discretize(MVPAL[,9]))
#mutinformation(discretize(MVPAL[,8]),discretize(MVPAL[,10]))
#mutinformation(discretize(MVPAL[,8]),discretize(MVPAL[,11]))
#mutinformation(discretize(MVPAL[,8]),discretize(MVPAL[,12]))
#mutinformation(discretize(MVPAL[,8]),discretize(MVPAL[,13]))
#mutinformation(discretize(MVPAL[,8]),discretize(MVPAL[,14]))
#mutinformation(discretize(MVPAL[,8]),discretize(MVPAL[,15]))
#mutinformation(discretize(MVPAL[,8]),discretize(MVPAL[,16]))
#mutinformation(discretize(MVPAL[,8]),discretize(MVPAL[,17]))
#mutinformation(discretize(MVPAL[,8]),discretize(MVPAL[,18]))
#mutinformation(discretize(MVPAL[,8]),discretize(MVPAL[,19]))
#mutinformation(discretize(MVPAL[,8]),discretize(MVPAL[,20]))
#mutinformation(discretize(MVPAL[,8]),discretize(MVPAL[,21]))
```

#Test and Train on MLB
#Obtaining a random year to be our testing data to predict MVP shares
```{r}
randomYear <- sample(1956:2022,1)
```

```{r}
train.data <- subset(MVPMLB, Year!=randomYear)
test.data <- subset(MVPMLB, Year==randomYear)
testAL.data <- subset(MVPAL, Year==randomYear)
testNL.data <- subset(MVPNL, Year==randomYear)


MVPAL.lm <- MVPAL
MVPNL.lm <- MVPNL
```

#Linear Regression with all statistics 
```{r}
MVP.lm.fit <- lm(Share~WAR+SB+HR+RBI+BA+OBP, data=train.data)
```

```{r}
summary(MVP.lm.fit)
```
#VIF Test
```{r}
vif(MVP.lm.fit)
```
#Predicting AL
```{r}
predict(MVP.lm.fit,newdata=testAL.data)
```
#Prediciting NL
```{r}
predict(MVP.lm.fit,newdata=testNL.data)
```

```{r}
MVPAL.lm$predicted <- predict(MVP.lm.fit,newdata=MVPAL)
MVPNL.lm$predicted <- predict(MVP.lm.fit,newdata=MVPNL)

```

```{r}
MVPAL.lm.mod <- MVPAL.lm %>%
  slice_max(predicted,by=c(Year))
```

```{r}
MVPNL.lm.mod <- MVPNL.lm %>%
  slice_max(predicted,by=c(Year))
```

```{r}
sum(MVPAL.lm.mod$Winners)
sum(MVPNL.lm.mod$Winners)
```
```{r}
MVPMLB.lm.mod <- rbind(MVPAL.lm.mod,MVPNL.lm.mod)
MVPMLB.lm.mod$Winners <- MVPMLB.lm.mod$Winners.Predicted
MVPMLB.lm.mod$Winners.Precited <- 1

```

```{r}
MVPMLB.lm.mod$Winners.Actual <- ifelse(MVPMLB.lm.mod$Rank == 1,1,0)
```

```{r}
rmse(MVPMLB.lm.mod$Winners.Actual,MVPMLB.lm.mod$Winners.Precited)
```

#Durbin Watson Test for independence
```{r}
durbinWatsonTest(MVP.lm.fit)
```

#Homoscedasticity test with Breusch Pagan Test
```{r}
ols_test_breusch_pagan(MVP.lm.fit)
```

#Running a Q-Q Plot to check the normality
```{r}
ggplot() +
  geom_qq(aes(sample=rstandard(MVP.lm.fit))) +
  geom_abline(color="red") +
  coord_fixed()
```
#Predicting on 2023 Year
```{r}
Batting2023$Predict <-predict(MVP.lm.fit,newdata=Batting2023)

```

#For loop to run our model multiple times to see how accurate it is to predicting the actual winner.
```{r}
ALPredicted <- data.frame()
NLPredicted <- data.frame()
i <- 1
for (i in 1:10){
  MVP.lm.fit <- lm(Share~WAR+SB+R+H+HR+RBI+BA+OBP, data=train.data)
  Batting2023AL$Predictions <- predict(MVP.lm.fit,newdata=Batting2023AL)
  Batting2023AL <- Batting2023AL[order(-Batting2023AL$Predictions),]
  Batting2023NL$Predictions <- predict(MVP.lm.fit,newdata=Batting2023NL)
  Batting2023NL <- Batting2023NL[order(-Batting2023NL$Predictions),]
  
  outputAL <-head(Batting2023AL,1)
  ALPredicted <- rbind(ALPredicted,outputAL)
  
  outputNL <-head(Batting2023NL,1)
  NLPredicted <- rbind(NLPredicted,outputNL)
}
```

```{r}
ALPredicted
```

```{r}
NLPredicted
```

```{r}

```