---
title: "Untitled"
author: "Ethan Michaels"
date: "2024-07-06"
output: pdf_document
---
#Install Libraries
```{r}
library(dplyr)
library(tidyverse)
library(randomForest)
library(caret)
library(car)
library(infotheo)
```

#Reading in Datasets
```{r}
MVPAL <- read.csv("MVPALContenders.csv", header = TRUE)
MVPNL <- read.csv("MVPNLContenders.csv", header = TRUE)
MVPMLB <- read.csv("MVPMLBWinners.csv", header = TRUE)
```

#Assigning a Dummy Variable to assign the winners of each Years MVP
```{r}
MVPAL$Winners <- ifelse(MVPAL$Rank==1,1,0)
MVPNL$Winners <- ifelse(MVPNL$Rank==1,1,0)
MVPMLB$Winners <- ifelse(MVPMLB$Rank==1,1,0)
```

#Modifying the Share Variable to be decimal points 
```{r}
MVPAL$Share <- as.numeric(sub("%", "", MVPAL$Share,fixed=TRUE))/100
MVPNL$Share <- as.numeric(sub("%", "", MVPNL$Share,fixed=TRUE))/100
MVPMLB$Share <- as.numeric(sub("%", "", MVPMLB$Share,fixed=TRUE))/100
```

#Calculating Batting Average
```{r}
Batting2023$BA <- (Batting2023$H/Batting2023$AB)
```

#Removing all Observations with less than 150 AB (As this implies they are pitchers)
```{r}
MVPAL <- subset(MVPAL, AB>150)
MVPNL <- subset(MVPNL, AB>150)
MVPMLB <- subset(MVPMLB, AB>150)
```

#Merging the AL and NL Dataset together
```{r}
MVPALNL <- dplyr::bind_rows(MVPAL,MVPNL)
```

#Ordering the dataset by Year
```{r}
MVPMLB <- MVPMLB[order(MVPMLB$Year), ]
```

#Removing all Pitching related statistics as it doesn't apply
```{r}
MVPALContendersTest <- MVPAL[c(1:21,34)]

MVPNLContendersTest <- MVPNL[c(1:21,34)]

MVPALNLContendersTest <- MVPALNL[c(1:21,34)]

```

```{r}

```

```{r}

```

```{r}
write.csv(MVPALNLContendersTest,"MVPALNLContenders.csv")
```