---
title: "Untitled"
author: "Ethan Michaels"
date: "2024-08-08"
output: pdf_document
---
#Installing Librarys
```{r}
library(dplyr)
library(randomForest)
library(caret)
```

#Reading in the relevant files
```{r}
MVPMLB <- read.csv("MVPALNLContendersModified.csv", header = TRUE)
Batting2023 <- read.csv("BattingData2.csv", header = TRUE)
Batting2024 <- read.csv("Batting2024.csv", header = TRUE)
```

#Creating the batting average statistic for 2023 data and dividing the 2023 datasets between MLBs Two leagues
```{r}
Batting2023$BA <- Batting2023$H/Batting2023$AB
Batting2023AL <- subset(Batting2023,League=="AL")
Batting2023NL <- subset(Batting2023,League=="NL")
```

#Creating dataframes for the dataset for years 1956-2022
```{r}
MVPAL <- MVPMLB[MVPMLB$League=="AL",]
MVPNL <- MVPMLB[MVPMLB$League=="NL",]
MVPWinners <- MVPMLB[MVPMLB$Winners==1,]
MVPLosers<- MVPMLB[MVPMLB$Winners==0,]
```

#Test and Train on MLB
#Obtaining a random year to be our testing data to predict MVP shares
```{r}
randomYear <- sample(1956:2022,1)
```

```{r}
train.data <- subset(MVPMLB, Year!=randomYear)
test.data <- subset(MVPMLB, Year==randomYear)
testAL.data <- subset(MVPAL, Year==randomYear)
testNL.data <- subset(MVPNL, Year==randomYear)

eq1 = Share ~ WAR+SB+H+HR+RBI+BA
```

```{r}
control <- trainControl(method ="cv",number=10, search = "grid")
```

#Tuning the Grid
```{r}
tuneGrid <- expand.grid(mtry=c(2,3,4,5,6,7,8))
```

```{r}
set.seed(1234)
```

#Running the random forest model
```{r}
rf_model <-  randomForest(Share ~WAR+SB+H+HR+RBI+BA+OBP,data=train.data,method="rf",trControl=control, tuneGrid=tuneGrid, importance=TRUE )

```

```{r}
print(rf_model)
```
#Predicting on each MLB League
```{r}
predictionsAL <- predict(rf_model,newdata=MVPAL)
predictionsNL <- predict(rf_model,newdata=MVPNL)

```

#making the predictions into a dataframe
```{r}
predictionsAL <- data.frame(predictionsAL)
predictionsNL <- data.frame(predictionsNL)

```

#Inserting the predictions back into the dataset
```{r}
MVPAL$predicted <- predictionsAL
MVPNL$predicted <- predictionsNL

```

#Splicing the data to only display the predicted winner of each year for AL and NL
```{r}
MVPAL.rf.mod <- MVPAL %>%
  slice_max(predicted,by=c(Year))
```

```{r}
MVPNL.rf.mod <- MVPNL %>%
  slice_max(predicted,by=c(Year))
```

#Below is code to be able to run a confusion matrix
```{r}
MVPAL.rf.mod$Predicted_Winners <- 1
MVPNL.rf.mod$Predicted_Winners <- 1

#MVPAL.rf$Predicted_Winners<- MVPAL.rf.mod$Predicted_Winners
```

#Merging datasets together to display predicted winners and actual winners
```{r}
MVPAL$ID <- paste(MVPAL$Name,MVPAL$Year)
MVPAL.rf.mod$ID <- paste(MVPAL.rf.mod$Name,MVPAL.rf.mod$Year)
MVPAL.rf.mod<- MVPAL.rf.mod[,c("ID","Predicted_Winners")]

MVPNL$ID <- paste(MVPNL$Name,MVPNL$Year)
MVPNL.rf.mod$ID <- paste(MVPNL.rf.mod$Name,MVPNL.rf.mod$Year)
MVPNL.rf.mod<- MVPNL.rf.mod[,c("ID","Predicted_Winners")]
```

```{r}
MVPAL.testing <- left_join(MVPAL,MVPAL.rf.mod,by="ID")
MVPAL.testing$Predicted_Winners <- ifelse(is.na(MVPAL.testing$Predicted_Winners),0,1)

MVPNL.testing <- left_join(MVPNL,MVPNL.rf.mod,by="ID")
MVPNL.testing$Predicted_Winners <- ifelse(is.na(MVPNL.testing$Predicted_Winners),0,1)
```

#Making tables of the Predicted Winners and Actual winners to run a confusion matrix on
```{r}
MVPALtable <- table(MVPAL.testing$Winners,MVPAL.testing$Predicted_Winners)
MVPNLtable <- table(MVPNL.testing$Winners,MVPNL.testing$Predicted_Winners)
```

#Creating a confusion matrix for AL and NL
```{r}
MVPAL.cm <- caret::confusionMatrix(MVPALtable)
MVPAL.cm
```

```{r}
MVPNL.cm <- caret::confusionMatrix(MVPNLtable)
MVPNL.cm
```
#Finding the RMSE Value of AL and NL
```{r}
mean((MVPAL.testing$Predicted_Winners-MVPAL.testing$Winners)^2)
mean((MVPNL.testing$Predicted_Winners-MVPNL.testing$Winners)^2)

```

#Testing our model on 2023 data 
```{r}
Batting2023$Predictions <- predict(rf_model,newdata=Batting2023)
```

```{r}
Batting2023 <- Batting2023[order(-Batting2023$Predictions),]
```

```{r}
Batting2023AL$Predictions <- predict(rf_model,newdata=Batting2023AL)
```

```{r}
Batting2023AL <- Batting2023AL[order(-Batting2023AL$Predictions),]
```

```{r}
Batting2023NL$Predictions <- predict(rf_model,newdata=Batting2023NL)
```

```{r}
Batting2023NL <- Batting2023NL[order(-Batting2023NL$Predictions),]
```

```{r}
head(Batting2023AL,1)

```

```{r}
head(Batting2023NL,1)

```
#For loop to run our model multiple times to see how accurate it is to predicting the actual winner.
```{r}
ALPredicted <- data.frame()
NLPredicted <- data.frame()
i <- 1
for (i in 1:100){
  rf_model <-  randomForest(Share ~WAR+SB+H+HR+RBI+BA+OBP,data=train.data,importance=TRUE )
  Batting2023AL$Predictions <- predict(rf_model,newdata=Batting2023AL)
  Batting2023AL <- Batting2023AL[order(-Batting2023AL$Predictions),]
  
  Batting2023NL$Predictions <- predict(rf_model,newdata=Batting2023NL)
  Batting2023NL <- Batting2023NL[order(-Batting2023NL$Predictions),]
  
  outputAL <-head(Batting2023AL,1)
  ALPredicted <- rbind(ALPredicted,outputAL)
  
  outputNL <-head(Batting2023NL,1)
  NLPredicted <- rbind(NLPredicted,outputNL)
  }
```

#Displaying the List of AL Winners Predicted
```{r}
ALPredicted
```
#Displaying a Table of the occurances of each MVP Winner
```{r}
ALPredicted %>%
  group_by(NAME) %>%
  summarize(count = n())
```
#Displaying the List of NL MVP WInners Predicted
```{r}
NLPredicted
```
#Displaying a Table of the occurances of each MVP Winner
```{r}
NLPredicted %>%
  group_by(NAME) %>%
  summarize(count = n())
```
#Using this years batting stats predicting the future MVP Winners of the AL and NL.
```{r}
Batting2024AL <- subset(Batting2024,Lg == "AL")
Batting2024NL <- subset(Batting2024,Lg == "NL")

ALPredicted24 <- data.frame()
NLPredicted24 <- data.frame()
i <- 1
for (i in 1:100){
  rf_model <-  randomForest(Share ~WAR+SB+H+HR+RBI+BA+OBP,data=train.data,importance=TRUE )
  Batting2024AL$Predictions <- predict(rf_model,newdata=Batting2024AL)
  Batting2024AL <- Batting2024AL[order(-Batting2024AL$Predictions),]
  Batting2024NL$Predictions <- predict(rf_model,newdata=Batting2024NL)
  Batting2024NL <- Batting2024NL[order(-Batting2024NL$Predictions),]
  
  outputAL24 <-head(Batting2024AL,1)
  ALPredicted24 <- rbind(ALPredicted24,outputAL24)
  
  outputNL24 <-head(Batting2024NL,1)
  NLPredicted24 <- rbind(NLPredicted24,outputNL24)
  }

```

```{r}
ALPredicted24
```

```{r}
ALPredicted24 %>%
  group_by(Player) %>%
  summarize(count = n())
```

```{r}
NLPredicted24
```

```{r}
NLPredicted24 %>%
  group_by(Player) %>%
  summarize(count = n())
```